---
title       : Predicting Miles Per Gallon
subtitle    : Homework Assignment of Developing Data Products
author      : tucuxi
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
---

## Overview

**Fuel efficiency** is getting increasingly important for car buyers.

A popular measure for fuel efficiency is **Miles Per Gallon** (mpg).

Using R and Shiny, we build a **web application** for predicting mpg from basic
car specifications.

The presentation is organized as follows:

1. Data set
2. Linear model for predicting miles per gallon
3. Shiny application

--- 

## Data Set

```{r echo = FALSE}
data(mtcars)
```

We use the **mtcars** data set from the **datasets** package, which contains
data from Motor Trend US magazine for 32 cars. This will be the basis for the
linear model.

Column  | Variable    | Description
------- | ----------- | -------------------
1       | mpg         | Miles/(US) gallon
2       | cyl         | Number of cylinders
3       | disp        | Displacement (cu.in.)
4       | hp          | Gross horsepower
5       | drat        | Rear axle ratio
6       | wt          | Weight (lb/1000)
7       | qsec	      | 1/4 mile time
8       | vs	      | V/S
9       | am	      | Transmission (0 = automatic, 1 = manual)
10      | gear	      | Number of forward gears
11      | carb	      | Number of carburetors

---

## Linear Model for Predicting Miles Per Gallon

We build a **linear model** for mpg with two predictors:
weight (wt) and number of cylinders (cyl).

```{r echo = FALSE}
lm <- lm(mpg ~ wt + factor(cyl), data = mtcars)
print(lm)
```

This model has **statistically significant P-values** (< 0.05) as shown below:

```{r}
summary(lm)$coefficients[, 4]
```

---

## Shiny Application

The web application built with [Shiny](http://shiny.rstudio.com)
allows the user to enter the number of cylinders and weight.

It **plots the linear model** and
**highlights the predicted mpg value** in the diagram.

At the heart of the application is the calculation of the mpg value. The
value is computed according to the linear model by the **predict()**
function.

For instance, for a car with 6 cylinders and a weight of 3,500 lbs, our
model predicts:

```{r}
predict(lm, data.frame(cyl = 6, wt = 3.5))
```

Try out the application [here](http://tucuxi.shinyapps.io/data_products)!

<style>
strong {
  font-weight: bold;
}
</style>
